---
title: "R Notebook"
output: html_notebook
---

# package googledrive

Using the googledrive package
https://cran.r-project.org/web/packages/googledrive/googledrive.pdf

```{r}
##install googledrive package
#install.packages("googledrive")

#load package
library(googledrive)
```

Now that the package is loaded, let's work with an example of the IPDGC project tracker sheet.
We will use the as_dribble function. We can pass this function a file ID.

Where to get our file ID? See this stackopverflow answer:
https://stackoverflow.com/questions/15057907/how-to-get-the-file-id-so-i-can-perform-a-download-of-a-file-from-google-drive-a

"...The easiest and fastest way to get a Google Drive file ID is from Google Drive on the web. 
Right-click the file name and select Get shareable link. 
The last part of the link is the file ID. Then you can cancel the sharing."

This is the file ID for the example spreadsheet:
https://docs.google.com/spreadsheets/d/1YBUOskdNmd1ZLiNJP5F-OS58lN5vBhYtrzSF_9HWoz0/edit?usp=sharing

It looks like it prompts users to sign in with their google account.
```{r}

# specify the file id (substitute a real file id of your own!)
IPDGC_exsheet <- as_dribble(as_id("1YBUOskdNmd1ZLiNJP5F-OS58lN5vBhYtrzSF_9HWoz0"))

```

For the real spreadsheet, I believe we use the as_team_drive function
instead of as_dribble. Then users will sign in with their google account.
```{r}

## specify the id (substitute with real IPDGC spreadsheet file ID!)
as_team_drive(as_id("0AOPK1X2jaNckUk9PVA"))

```

Let's look at some utility functions.
```{r}

is_dribble(IPDGC_exsheet) #is this even a dribble?

is_mine(IPDGC_exsheet) #Is it mine?

is_team_drive(IPDGC_exsheet)  #is it a team drive




```

Get info on Drive and Drive user capabilities
```{r}
drive_user()
```

I don't think the googledrive package allows you to edit exisiting sheets on Google Drive. 
But there is a package called googlesheets4 which does this.

# package googlesheets4:

https://googlesheets4.tidyverse.org
```{r}
#install.packages("googlesheets4")
library(googlesheets4)
```

googlesheets4 is pipe-friendly (and reexports %>%), but works just fine without the pipe.

You can read googlesheets data from:

a URL
a Sheet ID
a dribble produced by the googledrive package, which can lookup by file name
```{r}

#Authorize googlesheets4 to view and manage your Google Sheets.
gs4_auth()

#we can read from dribble object, created with googledrive package
read_sheet(IPDGC_exsheet)

#or we can read by using the url to the spreadsheet
read_sheet("https://drive.google.com/open?id=1YBUOskdNmd1ZLiNJP5F-OS58lN5vBhYtrzSF_9HWoz0")
```
Note that we have an issue with this function because the actual column names dont begin until row 2.
An easy fix is getting rid of the "Current Projects" cell and moving the rest of the sheet up to row 1.
I think this is ideal, but here's another workaround
```{r}

my_sheet <- read_sheet(IPDGC_exsheet)

colnames(my_sheet) <- my_sheet[1, ]
colnames(my_sheet)

```
We can access the google sheet like a normal data.frame object
```{r}

dim(my_sheet) #dimensions

my_sheet$Status #acess the Status column

str(my_sheet)

#print out each entry from every column
sapply(colnames(my_sheet), function(x) my_sheet[, x])


#Subset to show only ongoing projects
subset(my_sheet, Status == "ongoing")

#since googlesheets is part of the tidyverse, we can also use pipes with the magrittr package
library(magrittr)

my_sheet %>% subset(Status == "ongoing")

#only show projects by a specific PI (e.g. Nalls)
my_sheet[ grepl("Nall", my_sheet$`Proposed by (PI)`, ignore.case = T),  ] #the grepl functions returns a boolean vector indicating whether that cell/line contains "Nalls"

```
## Opening our googlesheet in the broswer
```{r}
gs4_browse(IPDGC_exsheet)
```


## Editing cells

We can use the range_write function to edit our sheet
Let's edit cell A3 and change it from "ongoing" to "test"
Note that we need to wrap "test" inside as.data.frame

If you have the google sheet open in your broswer, you can see these edits in real time.
```{r}
#edit
range_write(IPDGC_exsheet, as.data.frame("test"), range = "A3")

#revert back to "ongoing"
range_write(IPDGC_exsheet, as.data.frame("ongoing"), range = "A3")


```


